---
title: "ABCD Multiview Step 1. Dataset, Simulation Topic, Questions, and Relevant References"
author: "Aidan Neher"
format: docx
editor: visual
---

## Motivating Questions

1. Motivating application: Which early life adversity (ELA), neuroimaging, and genomic features are important for the prediction of cognitive outcomes? 
2. Methods extension: Does extension of the BIP framework to a mixed outcome model that accommodates observational clustering perform well, specifically above and beyond the BIP framework that does not account for observational clustering, in this setting and related settings?  

## ABCD Study Dataset

The Adolescent Brain and Cognitive Development Study (ABCD Study®) is the largest long-term study of brain development and child health in the United States. Starting with students aged 9-10 years old and repeated until age 19-20, brains scans are taken bi-annually, biosamples, paper and pen tests, and iPad tasks annually, and interviews every 3-6 months.

We are most interested in the following "views"

* Early life adversity (ELA)
* Neuroimaging 
  + sMRI: Cortical Thickness (CT) and Surface Area (SA)
  + fMRI: Functional Connectivity (FC)
* Genomics/ epigenomics

We are performing variable selection jointly with prediction in relation to the following continuous outcomes

* internalizing problems, 
* externalizing problems, 
* cognitive flexibility, and 
* inhibitory control

We control for common covariates 
* sex
* age
* race
* ... 

Firstly, we load the ELA view using code shared by Mark Fiecas. Here is the additional context he provided on sharing: "This came from the supplementary material of the Orendain et al paper. Note that they actually used more, but these were the ones that were "relevant" per their factor analysis. They had about ~10 more variables but those did not have high factor loadings. Please see their supplemental material if you'd like to get the other ~10. You'll need to revise Ellery's standard prep code to make sure these get loaded in (see her "files" variable), and the first place to start is here, which is at the very bottom."

```{r,include=FALSE}

library(tidyverse)

data_dir <- "/Users/aidanneher/Library/CloudStorage/Box-Box/ABCD Tabulated Data/5.1/core"

# Defaults to filtering down to baseline data - alternatively, pass a vector of eventnames
load_ela_data <- function(data_dir, events = "baseline_year_1_arm_1") {
  
  # Physical and sexual violence
  mh_p_ksads_path <- file.path(data_dir, "mental-health/mh_p_ksads_ptsd.csv")
  mh_p_ksads <- read_csv(mh_p_ksads_path) %>%
    select(src_subject_id, eventname, ksads_ptsd_raw_761_p,
           ksads_ptsd_raw_762_p, ksads_ptsd_raw_763_p,
           ksads_ptsd_raw_767_p, ksads_ptsd_raw_768_p,
           ksads_ptsd_raw_769_p, ksads_ptsd_raw_760_p,
           ksads_ptsd_raw_764_p, ksads_ptsd_raw_765_p)
  
  # Parent psychopathology
  mh_p_fhx_path <- file.path(data_dir, "mental-health/mh_p_fhx.csv")
  mh_p_fhx <- read_csv(mh_p_fhx_path) %>%
    select(src_subject_id, eventname, famhx_4_p,
           fam_history_5_yes_no, fam_history_6_yes_no,
           fam_history_7_yes_no, fam_history_8_yes_no,
           fam_history_11_yes_no, fam_history_12_yes_no,
           fam_history_13_yes_no)
  
  # Neighborhood Threat
  ce_p_nsc_path <- file.path(data_dir, "culture-environment/ce_p_nsc.csv")
  ce_p_nsc <- read_csv(ce_p_nsc_path) %>%
    select(src_subject_id, eventname, neighborhood3r_p, neighborhood2r_p)
  
  # Prenatal Substance Exposure
  ph_p_dhx_path <- file.path(data_dir, "physical-health/ph_p_dhx.csv")
  ph_p_dhx <- read_csv(ph_p_dhx_path) %>%
    select(src_subject_id, eventname,
           devhx_9_tobacco, devhx_9_alcohol,
           devhx_9_marijuana, devhx_9_coc_crack,
           devhx_9_her_morph, devhx_9_oxycont)
  
  # Scarcity
  abcd_p_demo_path <- file.path(data_dir, "abcd-general/abcd_p_demo.csv")
  abcd_p_demo <- read_csv(abcd_p_demo_path) %>%
    select(src_subject_id, eventname, demo_fam_exp1_v2, demo_fam_exp5_v2)
  
  # Household Dysfunction
  ce_y_fes_path <- file.path(data_dir, "culture-environment/ce_y_fes.csv")
  ce_y_fes <- read_csv(ce_y_fes_path) %>%
    select(src_subject_id, eventname, fes_youth_q6, fes_youth_q1, fes_youth_q5)
  
  # Merging all data frames
  merged_data <- list(mh_p_ksads, mh_p_fhx, ce_p_nsc, ph_p_dhx, abcd_p_demo, ce_y_fes) %>%
    reduce(full_join, by = c("src_subject_id", "eventname"))
  
  # Filter to events of interest
  filtered_data <- merged_data %>%
    filter(eventname %in% events)
  
  # Recode values in all columns except 'src_subject_id' and 'eventname'
  tidy_data <- filtered_data %>%
    mutate(across(-c(src_subject_id, eventname), ~na_if(., 999))) %>%  # Recode 999 "Don't know" as NA
    mutate(across(-c(src_subject_id, eventname), ~na_if(., 7))) %>%    # Recode 7 "Refuse to answer" as NA
    mutate(across(-c(src_subject_id, eventname), ~na_if(., 777)))      # Recode 777 "Refuse to answer" as NA
  
  return(tidy_data)
}

ela_data <- load_ela_data(data_dir)

```

Let's look at our ELA data. 

```{r}

str(ela_data)

```

We check out the proportion of missingness in our ELA data.

```{r}

# Calculate the proportion of missing values per column, excluding specific columns
missing_data <- ela_data %>%
  select(-src_subject_id, -eventname) %>%  # Exclude these columns from the analysis
  summarise(across(everything(), ~sum(is.na(.))/n())) %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "prop_missing")

# Plot the proportion of missing data
ggplot(missing_data, aes(x = variable, y = prop_missing)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Variable", y = "Proportion Missing", title = "Proportion of Missing Data by ELA Variable at Baseline")

```

Additional things to look at: 

* tsne plots (works for binary or ordinal variables?) / PCA
* 0 variance in any variable? i.e. no observations for a given level
* Polychoric correlation structure

```{r}



```


Now we load the covariate and outcome data, adapting code from Ellery's "ABCD_data_prep.Rmd".

# NOTES FOR USERS: 

**Things you need to change:**
    1. In user input chunk, you must change the data directory and the output directory
    2. You must change the file path to load in the qc data set (line 91) 
    
**Things you may want to change:** 
    1. The time point list 
    2. If you want specific KSADs variables, other than ones related to depression, NSSI, and suicidal ideation, you must add those to the KSAD INITIAL CLEANING section
    3. If you want to keep "raw" variables (e.g., granular race questions, specific KSAD variable), you must comment out specific lines in the Raw Variable Removal section (begins at line 507)

**If you want other variables not captured in this document:**
    1. Find the file in which the variable(s) is stored and import that file using `read.csv`
    2. Use `select` in the dplyr package to choose the desired variables within that data set and make sure to select `src_subject_id`, and `eventname` so each observation is uniquely identified and so the eventual merging goes smoothly (see lines 229 - 284 for example)
    3. Add your variable to the list created in line 292 (this ensures that your variable is included in the merged data frame)
    4. Determine if your variable(s) need to be cleaned and perform necessary cleaning steps
    

# Code 

```{r user-input}
# Location of "ABCD Standard Data Preparation" Box Directory (on local computer)
# Note that example below is based on a Mac directory
data_dir <- "/Users/aidanneher/Library/CloudStorage/Box-Box/ABCD Tabulated Data/5.1/core"

# Location of desired output directory of created csv file - if NULL, will output file into current working directory
out_dir <- '/Users/aidanneher/Documents/GitHub/abcd_multiview/data'
# out_dir <- NULL

# Date you want used in output name of csv file - if NULL, will use output from Sys.Date() (current date)
# out_date <- '2023-03-03' 
out_date <- NULL

# Initials or other string you want used in output name of csv file - no NULL option here, have to put in some sort of string (though we could create a default for this?)
out_initials <- 'AN' 

# Do you also want to output a wide format data file?
wide <- TRUE # [TRUE,FALSE]

# What variables do you want to output in the wide format?
# If only a subset is desired, include string giving the name of each variable combined into a vector as c('var_1','var_2','var_3') 
# e.g., wide_vars <- c('eventname','demo_sex_v2','demo_gender_id_v2','demo_prnt_marital_v2')
# See codebook for variable labels: https://docs.google.com/spreadsheets/d/1uHRrXASaxtZbRAeJvPJktqxyDkPdA2_WgxYdLTjhK1Q/edit?usp=sharing
# If NULL, will use ALL variables included in the regular long output
wide_vars <- NULL 
# Note that this currently assumes the same time points are desired for long and wide formats - does this make sense?

# What time points do you want to output? Put the string provided for all desired timepoints into a vector using c('timepoint1_name','timepoint2_name') etc.
timepoint_list <- c("baseline_year_1_arm_1") #,"1_year_follow_up_y_arm_1", "2_year_follow_up_y_arm_1") 
```

USE ONLINE CODEBOOK TO FIND THE LABELS FOR THE VARIABLES: 
https://docs.google.com/spreadsheets/d/1uHRrXASaxtZbRAeJvPJktqxyDkPdA2_WgxYdLTjhK1Q/edit?usp=sharing

Load Data Files

```{r}

library(tidyverse)

# Define the base directory for the data
data_dir <- "/Users/aidanneher/Library/CloudStorage/Box-Box/ABCD Tabulated Data/5.1/core"

# Function to load the required datasets
load_datasets <- function(data_dir) {
  # Standard files for most studies
  demog <- read_csv(file.path(data_dir, "abcd-general/abcd_p_demo.csv"))
  demog_y <- read_csv(file.path(data_dir, "mental-health/mh_y_ksads_bg.csv"))
  puberty <- read_csv(file.path(data_dir, "physical-health/ph_p_pds.csv"))
  study_covars <- read_csv(file.path(data_dir, "abcd-general/abcd_y_lt.csv"))
  sib_twin <- read_csv(file.path(data_dir, "genetics/gen_y_pihat.csv"))
  
  # MRI standard files
  mri <- read_csv(file.path(data_dir, "imaging/mri_y_adm_info.csv"))
  qc <- read_csv(file.path(data_dir, "imaging/mri_y_qc_incl.csv"))
  scan_qtns <- read_csv(file.path(data_dir, "imaging/mri_y_adm_qtn.csv"))
  
  # Standard files for clinical data
  cbcl <- read_csv(file.path(data_dir, "mental-health/mh_p_cbcl.csv"))
  bpm_y <- read_csv(file.path(data_dir, "mental-health/mh_y_bpm.csv"))
  ksad_p <- read_csv(file.path(data_dir, "mental-health/mh_p_ksads_ss.csv"))
  ksad_y <- read_csv(file.path(data_dir, "mental-health/mh_y_ksads_ss.csv"))
  
  # Combine all datasets into a list (if needed for further processing or return)
  datasets <- list(demog = demog, demog_y = demog_y, puberty = puberty,
                   study_covars = study_covars, sib_twin = sib_twin, mri = mri,
                   qc = qc, scan_qtns = scan_qtns, cbcl = cbcl, bpm_y = bpm_y,
                   ksad_p = ksad_p, ksad_y = ksad_y)
  
  return(datasets)
}

# Load the datasets
datasets <- load_datasets(data_dir)

# Make data.frames accessible to global environment 
attach(datasets)

```


```{r}
# KSAD INITIAL CLEANING
# why: During the data collection, a new version of the KSAD was used (KSAD2), so each KSAD variable has a corresponding KSAD2 variable which contains the data since the new version was adopted. The code below combines these variables into one.

ksad_new_df <- ksad_p %>% # initialize new data frame
  select(src_subject_id, eventname)
  
clean_ksads <- function(ksad_new_df, ksad, ksad2, p_or_t){ 
  # goal: combine two ksad variables into one new var
  # ksad_new_df = a new data frame created above to hold the new vars, 
  # ksad = 1st ksad variable (old ksad), 
  # ksad2 = ksad2 var (same content/question as ksad1), 
  # p_or_t = whether the variable is a p or t ksad 
  # output: a dataframe containing the new variable
  assign("ksad", paste0(ksad, sep = "_", p_or_t)) # create ksad variable from inputs and assign it the name ksad
  assign("ksad2", paste0(ksad2, sep = "_", p_or_t)) # create ksad2 variable from inputs and assign it the name ksad2
  assign("ksad_new", paste0(ksad, "_new")) # create new variable name from inputs and assign it the name ksad_new
  assign("ksad_df", paste("ksad", if_else(p_or_t == "p", "p", "y"), sep = "_")) # create data frame name from inputs and name it ksad_df
  my_cols <- c("src_subject_id", "eventname", ksad, ksad2) # create vector of necessary column names
  intermediary <- get(ksad_df) %>%
    select(all_of(my_cols)) %>% # select cols (need to do it this way because I'm referring to the cols as strings)
    mutate(!!ksad_new := if_else(is.na(get(ksad)) == T, get(ksad2), get(ksad))) # create new var
  ksad_new_df <- right_join(intermediary, ksad_new_df, by = c("src_subject_id", "eventname")) # join new var to existing data set
  ksad_new_df <- ksad_new_df %>%
    select(src_subject_id, eventname, ends_with("_new")) # select only necessary vars
  return(ksad_new_df)
}


# call the function with each pair of ksad and ksad2 variables (make sure to assign the output to the same name so you create one data set with all the new variables)

ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_946", "ksads2_23_906", "p") # suicidal ideation
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_947", "ksads2_23_907", "p")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_948", "ksads2_23_908", "p")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_949", "ksads2_23_909", "p")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_950", "ksads2_23_910", "p")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_951", "ksads2_23_911", "p")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_957", "ksads2_23_917", "p")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_958", "ksads2_23_918", "p")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_959", "ksads2_23_919", "p")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_960", "ksads2_23_920", "p") 
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_961", "ksads2_23_921", "p")

ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_954", "ksads2_23_914", "p") # suicidal attempts, prep
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_965", "ksads2_23_925", "p")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_962", "ksads2_23_922", "p")

ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_966", "ksads2_23_926", "p") # NO suicidal ideation/behaviors
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_955", "ksads2_23_915", "p")

ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_143", "ksads2_23_134", "p") # non-suicidal self injury
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_144", "ksads2_23_135", "p")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_956", "ksads2_23_916", "p")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_945", "ksads2_23_905", "p")

ksad_new_df <- clean_ksads(ksad_new_df, "ksads_1_840",	"ksads2_1_790", "p") # depression
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_1_841",	"ksads2_1_791", "p")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_1_842",  "ksads2_1_792", "p")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_1_843",	"ksads2_1_793", "p")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_1_844",	"ksads2_1_794", "p")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_1_845",	"ksads2_1_795", "p")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_1_846",	"ksads2_1_796", "p")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_1_847",	"ksads_1_847", "p") # this one does not have a corresponding ksad2 (I put it here so it would still appear in the df)

# repeat function calls with t instead of p

ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_946", "ksads2_23_906", "t")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_947", "ksads2_23_907", "t")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_948", "ksads2_23_908", "t")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_949", "ksads2_23_909", "t")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_950", "ksads2_23_910", "t")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_957", "ksads2_23_917", "t")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_958", "ksads2_23_918", "t")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_959", "ksads2_23_919", "t")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_960", "ksads2_23_920", "t")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_961", "ksads2_23_921", "t")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_954", "ksads2_23_914", "t")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_965", "ksads2_23_925", "t")

ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_966", "ksads2_23_926", "t")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_955", "ksads2_23_915", "t")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_951", "ksads2_23_911", "t")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_962", "ksads2_23_922", "t")

ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_143", "ksads2_23_134", "t")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_144", "ksads2_23_135", "t")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_956", "ksads2_23_916", "t")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_945", "ksads2_23_905", "t")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_963", "ksads2_23_923", "t")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_964", "ksads2_23_924", "t")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_953", "ksads2_23_913", "t")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_23_952", "ksads2_23_912", "t")

ksad_new_df <- clean_ksads(ksad_new_df, "ksads_1_840",	"ksads2_1_790", "t") 
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_1_841",	"ksads2_1_791", "t")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_1_842",  "ksads2_1_792", "t")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_1_843",	"ksads2_1_793", "t")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_1_844",	"ksads2_1_794", "t")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_1_845",	"ksads2_1_795", "t")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_1_846",	"ksads2_1_796", "t")
ksad_new_df <- clean_ksads(ksad_new_df, "ksads_1_847",	"ksads_1_847", "t") # this one does not have a corresponding ksad2 (I put it here so it would still appear in the df)
```

Demographic Cleaning Pre-Merge

```{r}
# COMBINE BASELINE VARIABLES WITH LONGITUDINAL VARIABLES
# how it combines: if the observation occurred in year1/arm1 (the baseline) the combination variable takes the value of the baseline variable, if not the combo variable takes the value of the longitudinal variable

var_vect <- c("demo_brthdat_v2", "demo_gender_id_v2", "demo_prnt_ed_v2", "demo_prtnr_ed_v2" , "demo_prnt_marital_v2", "demo_comb_income_v2", "demo_roster_v2", "demo_fam_exp1_v2", "demo_fam_exp2_v2", "demo_fam_exp3_v2", "demo_fam_exp4_v2", "demo_fam_exp5_v2", "demo_fam_exp6_v2", "demo_fam_exp7_v2") # input any baseline variable name you wish to combine with long. var in this vector 


combine.vars <- function(var){ # takes baseline variable name (a string) and makes a new variable using the corresponding long. var, outputs demog dataset with new combo variables (combo variables will be named the baseline variable name with "_comb" at the end)
  assign("var_comb", paste0(var, "_comb"))
  assign("var_l", paste0(var, "_l"))
  demog <- demog %>%
    mutate(!!var_comb := if_else(eventname == "baseline_year_1_arm_1", get(var), get(var_l))) # if observation occurred in year 1, arm 1 use bl var if not use long var
  return(demog)
}

for (i in 1:length(var_vect)) { # loops through vector of all variables to be combined and applies the combine.vars function
  demog <- combine.vars(var_vect[i])}


```

Select Relevant Variables
```{r}
demog_bl <- demog[demog$eventname == "baseline_year_1_arm_1",]

demog_bl <- demog_bl %>% 
  select(src_subject_id, demo_sex_v2, demo_race_a_p___10,
         demo_race_a_p___11,demo_race_a_p___12, demo_race_a_p___13, demo_race_a_p___14,
         demo_race_a_p___15, demo_race_a_p___16, demo_race_a_p___17, demo_race_a_p___18,
         demo_race_a_p___19, demo_race_a_p___20, demo_race_a_p___21, demo_race_a_p___22,
         demo_race_a_p___23, demo_race_a_p___24, demo_race_a_p___25,demo_race_a_p___77, 
         demo_race_a_p___99, demo_ethn_v2, demo_prnt_marital_v2, demo_prnt_ed_v2, demo_prtnr_ed_v2, demo_comb_income_v2)

demog_bl <- dplyr::rename(demog_bl, demo_prnt_marital_v2_bl = demo_prnt_marital_v2)
demog_bl <- dplyr::rename(demog_bl, demo_comb_income_v2_bl = demo_comb_income_v2)
demog_bl <- dplyr::rename(demog_bl, demo_prnt_ed_v2_bl = demo_prnt_ed_v2)
demog_bl <- dplyr::rename(demog_bl, demo_prtnr_ed_v2_bl = demo_prtnr_ed_v2)

demog <- demog %>%
  select(src_subject_id, eventname, demo_brthdat_v2_comb, demo_gender_id_v2_comb, 
         demo_prnt_ed_v2_comb, demo_prtnr_ed_v2_comb, demo_prnt_marital_v2_comb, demo_comb_income_v2_comb,
         demo_roster_v2_comb, demo_fam_exp1_v2_comb, demo_fam_exp2_v2_comb, demo_fam_exp3_v2_comb, 
         demo_fam_exp4_v2_comb, demo_fam_exp5_v2_comb, demo_fam_exp6_v2_comb, demo_fam_exp7_v2_comb,
         acs_raked_propensity_score)

demog_y <- demog_y %>%
  select(src_subject_id, eventname, kbi_gender, kbi_y_trans_id, kbi_y_sex_orient)

puberty <- puberty %>%
  select(src_subject_id, eventname, pds_p_ss_male_category_2, pds_p_ss_female_category_2)

fam <- study_covars[study_covars$eventname == "baseline_year_1_arm_1",]

fam <- fam %>%
  select(src_subject_id, rel_family_id)

study_covars <- study_covars %>%
  select(src_subject_id, eventname, site_id_l, interview_age)

sib_twin <- sib_twin %>% 
  select(src_subject_id, rel_relationship, rel_group_id)

mri <- mri %>%
  select(src_subject_id, eventname, mri_info_manufacturer)

qc <- qc %>%
  select(src_subject_id, eventname, imgincl_rsfmri_include)

cbcl <- cbcl %>%
  select(src_subject_id, eventname, cbcl_scr_syn_internal_r, cbcl_scr_syn_external_r, cbcl_scr_syn_totprob_r,
         cbcl_scr_dsm5_depress_r, cbcl_scr_dsm5_anxdisord_r, cbcl_scr_dsm5_adhd_r,
         cbcl_scr_syn_internal_t, cbcl_scr_syn_external_t, cbcl_scr_syn_totprob_t,
         cbcl_scr_dsm5_depress_t, cbcl_scr_dsm5_anxdisord_t, cbcl_scr_dsm5_adhd_t)

bpm_y <- bpm_y %>% 
  select(src_subject_id, eventname, bpm_y_scr_attention_r, bpm_y_scr_attention_t, bpm_y_scr_internal_r, 
         bpm_y_scr_internal_t, bpm_y_scr_external_r, bpm_y_scr_external_t, bpm_y_scr_totalprob_r, 
         bpm_y_scr_totalprob_t)

```

Merge into single file and replace missing data codes with NA (Can change this if you do not want that)

```{r}

files <- list(demog, demog_y, puberty, study_covars, mri, qc, scan_qtns, cbcl, bpm_y, ksad_new_df)
abcd_data_0 <- files %>% reduce(full_join, by = c("src_subject_id", "eventname"))

files_2 <- list(demog_bl, fam, sib_twin, abcd_data_0)
abcd_data <- files_2 %>% reduce(full_join, by = "src_subject_id")

# AW 8/3/2023: I added another step for merging to feed forward the BL data. 

# MF: 3/10/23 Check if the lines below yield the same intended result. This bypasses the problem with the dplyr update
# EI: Where is the original code that replaces the missing data codes with NA? 
## AW: 9/27/23 We deleted it at some point but have been using Mark's code for this purpsoe for quite a while
abcd_data <- abcd_data %>%
  mutate(across(where(is.numeric), ~na_if(.,777))) %>%
  mutate(across(where(is.numeric), ~na_if(.,999))) %>%
  mutate(across(where(is.numeric), ~na_if(.,555))) %>%
  mutate(across(where(is.numeric), ~na_if(.,888))) %>%
  mutate(across(where(is.character), ~na_if(.,"777"))) %>%
  mutate(across(where(is.character), ~na_if(.,"999"))) %>%
  mutate(across(where(is.character), ~na_if(.,"555"))) %>%
  mutate(across(where(is.character), ~na_if(.,"888"))) %>%
  mutate(across(where(is.character), ~na_if(.,""))) 

```

Clean and Recode Age

```{r}

abcd_data <- abcd_data %>%
  mutate(demo_brthdat_v2_comb_clean = if_else(demo_brthdat_v2_comb > 21, 
                                              demo_brthdat_v2_comb/12, demo_brthdat_v2_comb), # convert months to years
         demo_brthdat_v2_comb_clean = if_else(demo_brthdat_v2_comb_clean < 8, NA, demo_brthdat_v2_comb_clean), # younger than 8 --> NA
         demo_brthdat_v2_comb_clean = trunc(demo_brthdat_v2_comb_clean)) %>% # remove decimals 
  mutate(interview_age_b = interview_age / 12) # convert months to years

```


Recode Race Variables
```{r}

abcd_data <- abcd_data %>%
  mutate(demo_ethn_v2 = abs(demo_ethn_v2 - 2), # change "2" to 0 to match other vars 
    White_race = demo_race_a_p___10,
         Black_race = demo_race_a_p___11,
         AIAN_race = if_else(rowSums(select(., num_range("demo_race_a_p___", 12:13))) >=1, 1, 0),
         NHPI_race = if_else(rowSums(select(., num_range("demo_race_a_p___", 14:17))) >=1, 1, 0),
         Asian_race = if_else(rowSums(select(., num_range("demo_race_a_p___", 18:24))) >=1, 1, 0),
         Other_race = demo_race_a_p___25,
         Missing_race = if_else(demo_race_a_p___99 == 1, 1, demo_race_a_p___77), 
         Missing_race = if_else(White_race == 1 | # if participant did not endorse any of these, mark as missing
                                Black_race == 1 | 
                                AIAN_race == 1 |
                                NHPI_race == 1 |
                                Asian_race == 1 | 
                                Other_race == 1 |
                                demo_ethn_v2 == 1, 0, 1),
        Missing_race = if_else(is.na(Missing_race) == T, 1, Missing_race)) %>% # mark all NAs in missing race as 1
   mutate(Indigenous_race = if_else(AIAN_race == 1 | NHPI_race == 1, 1, 0)) 
 
```

Recode Puberty Variable

```{r}

abcd_data <- abcd_data %>%
  mutate(pubertal_status = if_else(demo_sex_v2 == '1' | demo_sex_v2 == '3',
                                  pds_p_ss_male_category_2, # Set to this if the condition above is TRUE
                                  pds_p_ss_female_category_2)) # Otherwise set to this 

```

Recode Parent-Related Demographics
```{r}

abcd_data <- abcd_data %>%
  mutate(across(c(demo_prnt_ed_v2_comb, demo_prtnr_ed_v2_comb, demo_prnt_ed_v2_bl, demo_prtnr_ed_v2_bl), ~as.integer(.x))) %>%
  mutate(highest_demo_ed_comb = case_when(
                                    is.na(demo_prtnr_ed_v2_comb) == T ~ demo_prnt_ed_v2_comb,
                                    demo_prnt_ed_v2_comb > demo_prtnr_ed_v2_comb ~ demo_prnt_ed_v2_comb,
                                    demo_prnt_ed_v2_comb <= demo_prtnr_ed_v2_comb ~ demo_prtnr_ed_v2_comb),
         highest_demo_ed_bl = case_when(
                                    is.na(demo_prtnr_ed_v2_bl) == T ~ demo_prnt_ed_v2_bl,
                                    demo_prnt_ed_v2_bl > demo_prtnr_ed_v2_bl ~ demo_prnt_ed_v2_bl,
                                    demo_prnt_ed_v2_bl <= demo_prtnr_ed_v2_bl ~ demo_prtnr_ed_v2_bl))

```


Create STB variables
```{r}

abcd_data <- abcd_data %>%
  # aggregate ksad questions to create meaningful, new variables 
       mutate(SI_ever_y = case_when(
                                ksads_23_946_t_new + ksads_23_947_t_new + ksads_23_948_t_new + ksads_23_949_t_new + ksads_23_950_t_new > 0 ~ "present",
                                ksads_23_957_t_new + ksads_23_958_t_new + ksads_23_959_t_new + ksads_23_960_t_new + ksads_23_961_t_new > 0 ~ "past",
                                is.na(ksads_23_946_t_new) == T |is.na(ksads_23_947_t_new) == T |is.na(ksads_23_948_t_new) == T | 
                                  is.na(ksads_23_949_t_new) == T |is.na(ksads_23_950_t_new) == T |is.na(ksads_23_957_t_new) == T |
                                  is.na(ksads_23_958_t_new) == T |is.na(ksads_23_959_t_new) == T |is.na(ksads_23_960_t_new) == T |
                                  is.na(ksads_23_961_t_new) == T~ NA, # check if vars have NA and preserve the NAs 
                                TRUE ~ "never"), # TRUE is case_when's "else"  
              SI_ever_p = case_when( 
                                ksads_23_946_p_new + ksads_23_947_p_new + ksads_23_948_p_new + ksads_23_949_p_new + ksads_23_950_p_new > 0 ~ "present", 
	                              ksads_23_957_p_new + ksads_23_958_p_new + ksads_23_959_p_new + ksads_23_960_p_new + ksads_23_961_p_new > 0 ~ "past",
                                is.na(ksads_23_946_p_new) == T |is.na(ksads_23_947_p_new) == T |is.na(ksads_23_948_p_new) == T |
                                  is.na(ksads_23_949_p_new) == T |is.na(ksads_23_950_p_new) == T |is.na(ksads_23_957_p_new) == T |
                                  is.na(ksads_23_958_p_new) == T |is.na(ksads_23_959_p_new) == T |is.na(ksads_23_960_p_new) == T |
                                  is.na(ksads_23_961_p_new) == T~ NA,
	                              TRUE ~ "never"),
              SI_ever_p_y = case_when(
                                SI_ever_p == "present" | SI_ever_y == "present" ~ "present", 
                                SI_ever_p == "past" | SI_ever_y == "past" ~ "past",
                                SI_ever_p == "never" | SI_ever_y == "never" ~ "never"),
              SA_ever_y = case_when(
                                is.na(ksads_23_954_t_new) == T | is.na(ksads_23_965_t_new) == T ~ NA,
                                ksads_23_954_t_new > 0 ~ "present", 
                                ksads_23_965_t_new > 0 ~ "past",
                                ksads_23_954_t_new <= 0 & ksads_23_965_t_new <= 0 ~ "never"),
              SA_ever_p = case_when(
                                ksads_23_954_p_new > 0 ~ "present",
                                ksads_23_965_p_new > 0 ~ "past",
                                ksads_23_954_p_new <= 0 & ksads_23_965_p_new <= 0 ~ "never"),
              SA_ever_p_y = case_when(
                                SA_ever_p == "present" | SA_ever_y == "present" ~ "present",
                                SA_ever_p == "past" | SA_ever_y == "past" ~ "past",
                                SA_ever_p == "never" | SA_ever_y == "never" ~ "never"),
              STB_highest_ever_y = case_when(
                      	        is.na(ksads_23_954_t_new) == T | is.na(ksads_23_965_t_new) == T |is.na(ksads_23_963_t_new) == T| 
                      	          is.na(ksads_23_964_t_new) == T| is.na(ksads_23_953_t_new)== T | is.na(ksads_23_952_t_new) == T|
                      	          is.na(ksads_23_949_t_new) == T| is.na(ksads_23_960_t_new) == T|is.na(ksads_23_959_t_new) == T|
                      	          is.na(ksads_23_961_t_new) == T| is.na(ksads_23_948_t_new) == T| is.na(ksads_23_950_t_new) == T|
                      	          is.na(ksads_23_947_t_new) == T| is.na(ksads_23_958_t_new) == T| is.na(ksads_23_957_t_new) == T|
                      	          is.na(ksads_23_946_t_new)== T ~ NA, 
	                 
    	                           0 < (ksads_23_954_t_new) + (ksads_23_965_t_new) ~ "SA",
    	                           0 < (ksads_23_963_t_new) + (ksads_23_964_t_new) + (ksads_23_953_t_new) + (ksads_23_952_t_new) ~ "SA interrup/aborted",
    	                           0 < (ksads_23_949_t_new) + (ksads_23_960_t_new) ~ "SI intent",
    	                           0 < (ksads_23_959_t_new) + (ksads_23_961_t_new) + (ksads_23_948_t_new) + (ksads_23_950_t_new) ~ "SI plan/method",
    	                           0 < (ksads_23_947_t_new) + (ksads_23_958_t_new) ~ "SI active nonspecific",
    	                           0 < (ksads_23_957_t_new) + (ksads_23_946_t_new) ~ "SI passive",
    	                           TRUE ~ "none"),
	           STB_highest_current_y = case_when(
	                               is.na(ksads_23_954_t_new) == T | is.na(ksads_23_953_t_new) == T | is.na(ksads_23_952_t_new) == T|
	                                 is.na(ksads_23_949_t_new) == T| is.na(ksads_23_948_t_new) == T| is.na(ksads_23_950_t_new)== T |
	                                 is.na(ksads_23_947_t_new)== T | is.na(ksads_23_946_t_new)== T ~ NA,
	                                          
	                                0 < (ksads_23_954_t_new) ~ "SA",
	                                0 < (ksads_23_953_t_new) + (ksads_23_952_t_new) ~ "SA interrup/aborted", # in original version of code this level is skipped (all participants in this level became NA)-- in new version it's preserved
	                                0 < (ksads_23_949_t_new) ~ "SI intent",
	                                0 < (ksads_23_948_t_new) + (ksads_23_950_t_new) ~ "SI plan/method",
	                                0 < (ksads_23_947_t_new) ~ "SI active nonspecific",
	                                0 < (ksads_23_946_t_new) ~ "SI passive",
	                                TRUE ~ "none")) %>%
    mutate(across(c(SI_ever_y, SI_ever_p,SI_ever_p_y, SA_ever_y,SA_ever_p,SA_ever_p_y, STB_highest_ever_y,STB_highest_current_y), ~as.factor(.x))) # convert all new vars to factors

```

Create NSSI Variables 

```{r}
abcd_data <- abcd_data %>%
  mutate(NSSI_ever_y = case_when(
                            is.na(ksads_23_945_t_new) == T | is.na(ksads_23_956_t_new) == T ~ NA,
                            ksads_23_945_t_new > 0 ~ "present",
                            ksads_23_956_t_new > 0 ~ "past",
                            TRUE ~ "never"),
         NSSI_ever_p = case_when(
                            is.na(ksads_23_945_p_new) == T | is.na(ksads_23_956_p_new) == T ~ NA,
                            ksads_23_945_p_new > 0 ~ "present",
                            ksads_23_956_p_new > 0 ~ "past",  
                            TRUE ~"never"),
         NSSI_ever_p_y = case_when(
                            NSSI_ever_p == "present" | NSSI_ever_y == "present" ~ "present",
                            NSSI_ever_p == "past" | NSSI_ever_y == "past" ~ "past",
                            NSSI_ever_p == "never" | NSSI_ever_y == "never" ~ "never"), 
         SITB_ever_y = case_when(
                            is.na(NSSI_ever_y) == T | is.na(STB_highest_ever_y) == T ~ NA,
                            STB_highest_ever_y != "none" | NSSI_ever_y != "never" ~ 1,
                            TRUE ~ 0)) %>%
    mutate(across(c(NSSI_ever_y, NSSI_ever_p, NSSI_ever_p_y, SITB_ever_y), ~as.factor(.x)))

```

Create Depression Variables

```{r}
abcd_data <- abcd_data %>%
  mutate(MDD_ever_y = case_when(
                          is.na(ksads_1_840_t_new) == T |is.na(ksads_1_841_t_new) == T | is.na(ksads_1_842_t_new) == T ~ NA,
                          0 < (ksads_1_840_t_new) ~ "present",
                          0 < (ksads_1_841_t_new) ~ "partial remission",
                          0< (ksads_1_842_t_new) ~ "past",
                          TRUE ~ "never"),
         AnyDD_ever_y = case_when(
                          is.na(ksads_1_840_t_new) == T |is.na(ksads_1_841_t_new) == T | is.na(ksads_1_842_t_new) == T |
                              is.na(ksads_1_843_t_new) == T | is.na(ksads_1_844_t_new) == T |is.na(ksads_1_845_t_new) == T | 
                              is.na(ksads_1_846_t_new) == T | is.na(ksads_1_847_t_new) == T ~ NA,
                          0 < (ksads_1_840_t_new) + (ksads_1_843_t_new) + (ksads_1_846_t_new) ~ "present",
                          0 < (ksads_1_841_t_new) + (ksads_1_844_t_new) ~ "partial remission",
                          0 < (ksads_1_842_t_new) + (ksads_1_845_t_new) + (ksads_1_847_t_new) ~ "past",
                          TRUE ~ "never"),
         MDD_ever_p = case_when(
                          is.na(ksads_1_840_p_new) == T |is.na(ksads_1_841_p_new) == T | is.na(ksads_1_842_p_new) == T ~ NA,
                          0 < (ksads_1_840_p_new) ~ "present",
                          0 < (ksads_1_841_p_new) ~ "partial remission",
                          0 < (ksads_1_842_p_new) ~ "past",
                          TRUE ~ "never"),
         AnyDD_ever_p = case_when(
                          is.na(ksads_1_840_p_new) == T |is.na(ksads_1_841_p_new) == T | is.na(ksads_1_842_p_new) == T |
                              is.na(ksads_1_843_p_new) == T | is.na(ksads_1_844_p_new) == T |is.na(ksads_1_845_p_new) == T | 
                              is.na(ksads_1_846_p_new) == T | is.na(ksads_1_847_p_new) == T ~ NA,
                          0 < (ksads_1_840_p_new) + (ksads_1_843_p_new) + (ksads_1_846_p_new) ~ "present",
                          0 < (ksads_1_841_p_new) + (ksads_1_844_p_new) ~ "partial remission",
                          0 < (ksads_1_842_p_new) + (ksads_1_845_p_new) + (ksads_1_847_p_new) ~ "past",
                          TRUE ~ "never"), 
         MDD_ever_p_y = case_when(
                          MDD_ever_p == "present" | MDD_ever_y == "present" ~ "present",
                          MDD_ever_p == "partial remission" | MDD_ever_y == "partial remission" ~ "partial remission",
                          MDD_ever_p == "past" | MDD_ever_y == "past" ~ "past",
                          MDD_ever_p == "never" | MDD_ever_y == "never" ~ "never"),
         AnyDD_ever_p_y = case_when(
                          AnyDD_ever_p == "present" | AnyDD_ever_y == "present" ~ "present",
                          AnyDD_ever_p == "partial remission" | AnyDD_ever_y == "partial remission" ~ "partial remission",
                          AnyDD_ever_p == "past" | AnyDD_ever_y == "past" ~ "past",
                          AnyDD_ever_p == "never" | AnyDD_ever_y == "never" ~ "never"
                          )) %>%
    mutate(across(c(MDD_ever_y, AnyDD_ever_y, MDD_ever_p, AnyDD_ever_y, MDD_ever_p_y, AnyDD_ever_p_y), ~as.factor(.x)))

```

Remove raw variables

```{r}
abcd_data <- abcd_data %>%
  select(-num_range("demo_race_a_p___", 12:25), # race vars
         -num_range("ksads_23_", 946:950, "_t_new"), # SI, SA, NSSI vars
         -num_range("ksads_23_", 952:954, "_t_new"),
         -num_range("ksads_23_", 957:961, "_t_new"),
         -num_range("ksads_23_", 963:965, "_t_new"),
         -num_range("ksads_23_", 946:950, "_p_new"),
         -ksads_23_954_p_new,
         -num_range("ksads_23_", 957:961, "_p_new"), 
         -num_range("ksads_1_", 840:847, "_t_new"), # depression vars
         -num_range("ksads_1_", 840:847, "_p_new"))

```


Remove the time points you do not want

```{r Remove unwanted events}
# abcd_data.selected_time <- abcd_data[which(abcd_data$eventname %in% timepoint_list),]
# tidyverse alternative to the above is below
abcd_data.selected_time <- abcd_data %>%
  filter(eventname %in% timepoint_list)

```


Write your own file to merge with your other variables of interest

```{r write dataset}

# Creates name based on date and initials/string passed in by user. If no date given, use the current date
if(is.null(out_date)){
  out_date <- Sys.Date()
}
# If no initials/string is given, throw an error that user must input one
if(is.null(out_initials)){
  stop("No input given for 'out_initials'. User must provide string input to create output file name.")
}
# Now output file based on name and out_dir given. Recall that if out_dir is NULL, it will write to current wd
csv_out_name <- paste0(out_dir,'/',out_date,'_',out_initials,'.csv')
write.csv(abcd_data.selected_time, csv_out_name, row.names = FALSE) 

```

Reformat to wide data if desired
```{r wide}
# Get desired variables to put into wide format based on user input
# If user input is NULL, just use ALL variables from abcd_data
if(is.null(wide_vars)){
  wide_vars <- colnames(abcd_data)
}

# Now only doing wide format if it was set to TRUE above. If set to FALSE, nothing will happen
if(wide){
  
  # Initialize list to store data frames for each time point in
  data_years <- vector("list", length(timepoint_list))
  
  # Now loop through time points
  # lapply solution
  data_years = lapply(timepoint_list,
                      function(x){
                        data_tp <- abcd_data %>% 
                          filter(eventname == x) %>%
                          dplyr::select(c(src_subject_id,all_of(wide_vars)))
                        names(data_tp) <- paste0(names(data_tp),"_",x)
                        data_tp <- data_tp %>% 
                          dplyr::rename(subjectkey = paste0('src_subject_id','_',x))
                        })
  
  # Now merge all together
  abcd_wide <- data_years %>% reduce(full_join, by = "subjectkey") %>% rename("src_subject_id" = "subjectkey")
  
  # Now output file based on initials and date input given (or for date, created in above chunk) and out_dir given
  # Recall that if out_dir is NULL, it will write to current wd
  csv_out_name_wide <- paste0(out_dir,'/',out_date,'_',out_initials,'_wide','.csv')
  write.csv(abcd_wide, csv_out_name_wide, row.names = FALSE) # KAD: 4/4/23 Added to suppress row names since they're usually unwanted
}

```

## Simulation Topic

## Relevant References

## Appendix: All code for this document consolidated

```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```
