---
title: Bayesian Longitudinal Integrative Analysis and Prediction with Application
  to the ABCD Study
author: "Aidan Neher"
date: "2023-04-05"
output: powerpoint_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Where are we going?

* Research Introduction
* Motivation
* ABCD Study
* Model Formulation
* Ponderings/ Next Steps

## NSF GRFP? 

## Intro: Bayesian Integrative Analysis and Prediction with Application to Atherosclerosis Cardiovascular Disease (Chekouo and Safo 2020)

* Authors integrate clinical, demographic, and multi-omics data to ID genetic factors involved in artherosclerosis cardiovascular disease (ASCVD).
* Frequentist joint integrative analysis and prediction methods exist e.g. Safo et al. 2019 combine CCA and linear discriminant analysis, "encourag[ing] predictors that are connected and behave similarly to be selected or neglected together”.
* Authors leverage *Bayesian* hierarchical factor analysis models for *variable selection* and *prediction*.

## Motivation: Extension to Longitudinal Data Analysis

* Adolescent Brain and Cognitive Development (ABCD) Study is an observational, longitudinal study of US children aged 9-10 at baseline across 21 study sites. 
* Neuroimaging data is collected every 2 years: baseline (T1) and year 2 (T2) neuroimaging data is available in 4th release.
* Developmental trajectories have lifelong implications.

## ABCD Study: Sample Demographics

We have selected subjects with neuroimaging data at T1 and T2. 

```{r}

```

## ABCD Study: Missingness

We plan to impute subjects with missing covariates/ outcomes.

```{r}

```

## ABCD Study: Clinical Outcome Trajectories

We are considering inclusion of a random intercept.

```{r}

```

## Model Formulation: Integrating Multiple Data Types

* Indices: $j$ feature, $t$ time 1,2, $l$ component, $i$ subject in $c$ family in $d$ site (highest level)
* Omics: $X^{(k)}_{ijt}=\Sigma_{l=1}^rU_{ilt}a_{ljt}^{(k)}+\epsilon_{ijt}^{(k)}$ where $U$ latent factor, $a$ *invariant* factor loading, $r$ number of components.
* Assumptions: 
  + $U$ is $MVN(0,I_r)$ and $corr(U_{il1}, U_{il2})=\rho_l$
  + $\epsilon_{ijt}^{(k)} \overset{\mathrm{iid}}{\sim} N(0,\sigma_{jtk}^2)$ and $cov(\epsilon_{ij1},\epsilon_{ij2})=\rho_{jk}$
  
## Model Formulation: Incorporating Longitudinal and Mixed Effects

* Indices: $j$ feature, $t$ time 1,2, $l$ component, $i$ subject in $c$ family in $d$ site (highest level)
* Response: $Y_{it}=\Sigma_{l=1}^rU_{ilt}\beta_{lt}+\alpha t+Z_i \gamma_i+\beta_{0cd}+R_{icd}+\epsilon_{it}$ where $\alpha$ fixed temporal effect, $Z\gamma$ covariates, $\epsilon_{it}\overset{\mathrm{iid}}{\sim} N(0,\sigma_{it}^2)$ 
* Level 2 Intercept: $\beta_{0cd}=\delta_{00c}+L_{0cd}$
* Level 3 Intercept: $\delta_{00d}=\lambda_{000}+G_{00d}$
* $R\sim N(0,\sigma^2)$, $L\sim N(0,\tau^2)$, $G\sim N(0,\nu^2)$

## Ponderings/ Next Steps

* How many time points is required for longitudinal data analysis? Is two time points enough? 
  + Potentially use a different data set?
* Are we asking a biomedically relevant question? Is the priority more so about understanding adolescent brain changes specifically? Is the inclusion of a clinical outcome ahead of ourselves?
* Implement cross-sectional model in R
* Translate slow modules to C++
* Impute missing covariates
* Add longitudinal factors
* Add random intercept effects
* Simulation studies
* Apply to ABCD Study in full

## Acknowledgements

* Mark Fiecas, Thierry Chekouo, and Sandra Safo for their mentorship and expertise. 
* Saonli Basu, Tianzhong Yang, other faculty, fellow students, and family.
* Interdisciplinary Biostatistics Training in Genetics and Genomics program, which is funded by NIH T32 Training Grant: T32 GM132063, 2020--2025