---
title: "ABCD Study CT and Demographic Data Exploratory Data Analysis"
author: "Aidan Neher"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## What is Exploratory Data Analysis?

Exploratory Data Analysis (EDA) is a structured approach for understanding your data that can be used for research question and hypothesis development. EDA's overall objective is to get insights to make better decisions. Sub-objectives include:

- Identify correlated variables.
- Identify and deal with outliers.
- Identify trends across time.
- Identify trends across space. 
- Uncover patterns related to the response variable of interest.
- Create research questions to explore or hypotheses to test.
- Identify possible new data sources.

### Set-Up Environment

The .RDS file loaded below was generated using the script "code/0_get_data.R". 

```{r,echo=FALSE}
library(tidyverse)
library(ggseg) # https://drmowinckels.io/blog/2021-03-14-new-ggseg-with-geom/
tidy_data <- readRDS("data/2023-02-16-tidy_data.RDS")
```

### About the Variables

* subjectkey is the subject's unique identifier.
* eventname is the data collection point for an observation (row of data). Note, interview_age is also available in months.
* Brain structure metrics cortical thickness (thick), sulcus depth (sulc), and surface area (area) are included. For more on the meaning of these metrics, see https://doi-org.ezp2.lib.umn.edu/10.1007%2Fs00429-015-1177-6

### Explore the data

From http://www.sthda.com/english/articles/31-principal-component-methods-in-r-practical-guide/118-principal-component-analysis-in-r-prcomp-vs-princomp/#prcomp-and-princomp-functions: 

"There are two general methods to perform PCA in R :

* Spectral decomposition which examines the covariances / correlations between variables
* Singular value decomposition which examines the covariances / correlations between individuals

The function princomp() uses the spectral decomposition approach. The functions prcomp() and PCA()[FactoMineR] use the singular value decomposition (SVD)."

```{r,echo=FALSE}
table(tidy_data$eventname)
split_data = split(tidy_data, f = tidy_data$eventname)
baseline_data = split_data$baseline_year_1_arm_1 %>% ungroup
baseline_smri = select(baseline_data, starts_with("smri"))
# https://www.datacamp.com/tutorial/pca-analysis-r
library(ggcorrplot)
library(FactoMineR)
library(factoextra)
# check for NAs
colSums(is.na(baseline_smri)) %>% sum
baseline_smri_normalized = scale(baseline_smri)
corr_matrix = cor(baseline_smri_normalized)
ggcorrplot(corr_matrix)
data_pca = princomp(corr_matrix)
# summary(data_pca)
data_pca$loadings[, 1:2]
fviz_eig(data_pca, addlabels = TRUE)
fviz_pca_var(data_pca, col.var = "black")
# contribution to first 2 components
fviz_cos2(data_pca, choice = "var", axes = 1:2) 
fviz_pca_var(data_pca, col.var = "cos2",
            gradient.cols = c("black", "orange", "green"),
            repel = TRUE)
# regression
pca_2 = prcomp(baseline_smri_normalized)
pca_2$loadings[, 1:2]
fviz_eig(pca_2, addlabels = TRUE)
fviz_pca_var(pca_2, col.var = "black")
# contribution to first 2 components
fviz_cos2(pca_2, choice = "var", axes = 1:2) 
fviz_pca_var(pca_2, col.var = "cos2",
            gradient.cols = c("black", "orange", "green"),
            repel = TRUE)

# plot loadings

```

## Questions

* There's more than one demo_data observation per subjectkey, why?
* What other covariates are required?
* Regarding time point determination for DSEM, do we want to use interview age or eventname?